\bf Visualising How Non-linear Dimension Reduction Warps Your Data

### 2024-01-13 paper.qmd
Regenerated Figure 4

```{r}
#| echo: false
# distance |> dplyr::filter(distance == max(distance))
# df_bin_centroids |> dplyr::filter(row_number() %in% c(10, 65))

a1 <- colour_long_edges(.data = distance, benchmark_value = 15, 
                        triangular_object = tr1_object, distance_col = distance) + 
  geom_point(size = 2, colour = "#33a02c") + 
  #ggtitle("(b)") + 
  xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
  theme_light() +
  #coord_equal() +
  theme(legend.position = "none", plot.title = element_text(size = 5, hjust = 0.5, vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()#change legend key width
  )

UMAP_data_with_hb_id_selected <- UMAP_data_with_hb_id |> dplyr::filter(hb_id %in% c(31, 239)) |>
  dplyr::mutate(select_point = if_else(hb_id == 31, "data1", "data2"))

a2 <- ggplot(data = hex_full_count_df, aes(x = x, y = y)) + 
  geom_polygon(color = "black", aes(group = polygon_id), fill = "#ffffff") +
  geom_point(data = UMAP_data_with_hb_id_selected, aes(x = UMAP1, y = UMAP2, colour = select_point), alpha = 0.5) +
  scale_color_manual(values=c("#e6550d", "#000000")) +
  geom_point(data = df_bin_centroids_all |> dplyr::filter(hexID %in% c(31, 239)), aes(x = x, y = y), colour = "#33a02c") +
  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
  theme_light() +
  theme(legend.position = "none", plot.title = element_text(size = 5, hjust = 0.5, vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()#change legend key width
  )

@@

df1 <- df_all |> 
  dplyr::filter(hb_id == "31") |>
  dplyr::select(tidyselect::starts_with("x")) |>
  dplyr::mutate(type = "data1") ## original dataset

df2 <- df_all |> 
  dplyr::filter(hb_id == "239") |>
  dplyr::select(tidyselect::starts_with("x")) |>
  dplyr::mutate(type = "data2") ## original dataset

df3 <- df_all |> 
  dplyr::filter(!(hb_id %in% c("31", "239"))) |>
  dplyr::select(tidyselect::starts_with("x")) |>
  dplyr::mutate(type = "data") ## original dataset

df_b <- df_bin |> 
  dplyr::filter(hb_id %in% df_bin_centroids$hexID) |>
  dplyr::select(-hb_id) |>
  dplyr::mutate(type = "model") ## Data with summarized mean

@@

distance_df_small_edges <- distance |> 
  dplyr::filter(from == 10) |>
  dplyr::filter(to == 65) |>
  head(1)

lg_high <- langevitour::langevitour(df_exe[1:(length(df_exe)-1)], group = df_exe$type, pointSize = 3, levelColors = c("#6a3d9a", "#e6550d", "#000000", "#33a02c"), lineFrom = distance_df_small_edges$from, lineTo = distance_df_small_edges$to, lineColors = "#e31a1c")

@@

```{r}
#| warning: false
#| echo: false
#| label: fig-wkhighD
#| fig-cap: How the 2D model lift into high dimensions? (a) visualize the points and the hexagonal bin centroids related 31st and 239th hexagons, (b) visualization of the edge connected the 31th and 239th hexagons (colored in red) in the triangular mesh. A video of tour animation is available at (need to add the link of lg_high).

>>>>

```{r}
#| echo: false
#| warning: false
data <- read_rds("data/s_curve/s_curve.rds")
training_data <- read_rds("data/s_curve/s_curve_training.rds")
UMAP_data <- read_rds(file = "data/s_curve/s_curve_umap.rds")

num_bins_x <- 2

# ## Data set with all possible centroids in the hexagonal grid
# 
# full_centroid_df <- generate_full_grid_centroids(df_bin_centroids)
# 
# ## Generate all coordinates of hexagons
# hex_grid <- full_hex_grid(full_centroid_df)
# 
# hex_full_count_df <- generate_full_grid_info(df_bin_centroids)
# 
# write_rds(hex_full_count_df, "data/s_curve/s_curve_hex_2.rds")

hex_full_count_df <- read_rds("data/s_curve/s_curve_hex_2.rds")

# Create the tibble with x and y coordinates
tr_df <- tibble::tibble(x = tr1_object$x, y = tr1_object$y)

benchmark_value <- 19.4

# Filter and label small and long edges
distance_df_small_edges <- distance |>
  dplyr::filter(distance < benchmark_value) |>
  dplyr::mutate(type = "small_edges")

distance_df_long_edges <- distance |>
  dplyr::filter(distance >= benchmark_value) |>
  dplyr::mutate(type = "long_edges")

# Combine small and long edges
distance_edges <- dplyr::bind_rows(distance_df_small_edges, distance_df_long_edges)

tr_from_to_df_coord_with_group <- merge(tr_from_to_df, distance_edges, by = c("from", "to"))

UMAP_data_with_hb_id_selected <- UMAP_data_with_hb_id |> dplyr::filter(hb_id %in% c(2, 15)) |>
  dplyr::mutate(select_point = if_else(hb_id == 2, "data1", "data2"))

a1 <- ggplot(tr_df |> dplyr::filter((y == min(tr_df$y)) | (y == max(tr_df$y))), aes(x = x, y = y)) +
  geom_segment(
    aes(x = x_from, y = y_from, xend = x_to, yend = y_to, color = type),
    data = tr_from_to_df_coord_with_group
  ) +
  geom_point() +
  coord_equal() +
  scale_colour_manual(values = c("#000000", "#f6e8c3"))  + 
  geom_point(size = 3, colour = "#33a02c") + 
  #ggtitle("(b)") + 
  xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
  theme_light() +
  #coord_equal() +
  theme(legend.position = "none", plot.title = element_text(size = 5, hjust = 0.5, vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()#change legend key width
  )

a2 <- ggplot(data = hex_full_count_df, aes(x = x, y = y)) + 
  geom_polygon(color = "#e7e1ef", aes(group = polygon_id), fill = "#ffffff") +
  geom_point(data = UMAP_data_with_hb_id_selected, aes(x = UMAP1, y = UMAP2, colour = select_point), alpha = 0.5) +
  scale_color_manual(values=c("#e6550d", "#000000")) +
  geom_point(data = df_bin_centroids_all |> dplyr::filter(hexID %in% c(2, 15)), aes(x = x, y = y), colour = "#33a02c", size = 3) +
  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
  theme_light() +
  theme(legend.position = "none", plot.title = element_text(size = 5, hjust = 0.5, vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()#change legend key width
  )

@@

df1 <- df_all |> 
  dplyr::filter(hb_id == "2") |>
  dplyr::select(tidyselect::starts_with("x")) |>
  dplyr::mutate(type = "data1") ## original dataset

df2 <- df_all |> 
  dplyr::filter(hb_id == "15") |>
  dplyr::select(tidyselect::starts_with("x")) |>
  dplyr::mutate(type = "data2") ## original dataset

df3 <- df_all |> 
  dplyr::filter(!(hb_id %in% c("2", "15"))) |>
  dplyr::select(tidyselect::starts_with("x")) |>
  dplyr::mutate(type = "data") ## original dataset

df_b <- df_bin |> 
  dplyr::filter(hb_id %in% df_bin_centroids$hexID) |>
  dplyr::filter(hb_id %in% c("2", "15")) |>
  dplyr::select(-hb_id) |>
  dplyr::mutate(type = "model") ## Data with summarized mean

@@

lg_high <- langevitour::langevitour(df_exe[1:(length(df_exe)-1)], group = df_exe$type, pointSize = 3, levelColors = c("#6a3d9a", "#e6550d", "#000000", "#33a02c"), lineFrom = 1, lineTo = 2, lineColors = "#e31a1c")

@@

```{r}
#| warning: false
#| echo: false
#| label: fig-wkhighD
#| fig-cap: How the 2D model lift into high dimensions? (a) visualize the points and the hexagonal bin centroids related 2nd and 15th hexagons, (b) visualization of the edge connected the 2nd and 15th hexagons (colored in red) in the triangular mesh. A video of tour animation is available at <https://www.youtube.com/watch?v=vBvM30Plt24>.

### 2024-01-13 paper.tex
Regenerated Figure 4

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, breakable, sharp corners, interior hidden, enhanced, boxrule=0pt, frame hidden]}{\end{tcolorbox}}\fi

@@

\caption{\label{fig-wkhighD}How the 2D model lift into high dimensions?
(a) visualize the points and the hexagonal bin centroids related 31st
and 239th hexagons, (b) visualization of the edge connected the 31th and
239th hexagons (colored in red) in the triangular mesh. A video of tour
animation is available at (need to add the link of lg\_high).}

>>>>

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, breakable, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, enhanced, frame hidden, sharp corners]}{\end{tcolorbox}}\fi

@@

\caption{\label{fig-wkhighD}How the 2D model lift into high dimensions?
(a) visualize the points and the hexagonal bin centroids related 2nd and
15th hexagons, (b) visualization of the edge connected the 2nd and 15th
hexagons (colored in red) in the triangular mesh. A video of tour
animation is available at
\url{https://www.youtube.com/watch?v=vBvM30Plt24}.}
